---
title: "p8105_hw1_wc2750"
author: "Weixi Chen"
date: "9/17/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
### Create a data frame
```{r}
# create variables
vec_sample = rnorm (10) # a random sample of size 10 from a standard Normal distribution
vec_logical = vec_sample > 0 # a logical vector indicating whether elements of the sample are greater then 0
vec_char = c ("Hi", "my", "name", "is", "Weixi", "Chen", "nice", "to", "meet", "you") # a character vector of length 10
vec_factor = factor(c("low", "low", "low", "medium", "medium", "medium", "medium", "high", "high", "high")) # a factor vector of length 10, with 3 different factor “levels”

# create a data frame using tibble()
library(tibble)
data_frame = tibble(
  vec_sample,
  vec_logical,
  vec_char,
  vec_factor
)
data_frame
```

### Apply as.numeric function
```{r eval=FALSE}
# apply as.numeric
as.numeric (vec_logical)
as.numeric (vec_char)
as.numeric (vec_factor)
```

### Convert variables
```{r collapse=TRUE}
as.numeric (vec_logical) * vec_sample
as.factor (vec_logical) * vec_sample
as.numeric (as.factor (vec_logical)) * vec_sample
```

## Problem 2
### Make a plot for "penguins" dataset
```{r}
# load "penguins" dataset
data("penguins", package = "palmerpenguins")
```

The "penguins" dataset shows the variables of `r names(penguins)`. The species include `r levels (factor(penguins$species))` from `r levels (factor(penguins$island))` islands. The data comes from year `r levels(factor(penguins$year))`. The bill length ranges from `r min(penguins$bill_length_mm, na.rm = TRUE)`mm to `r max(penguins$bill_length_mm, na.rm = TRUE)`mm. The bill depth ranges from `r min(penguins$bill_depth_mm, na.rm = TRUE)`mm to `r max(penguins$bill_depth_mm, na.rm = TRUE)`mm. The flipper length ranges from `r min(penguins$flipper_length_mm, na.rm = TRUE)`mm to `r max(penguins$flipper_length_mm, na.rm = TRUE)`mm. And the body mass ranges from `r min(penguins$body_mass_g, na.rm = TRUE)`g to `r max(penguins$body_mass_g, na.rm = TRUE)`g.

The dataset has `r nrow (penguins)` rows and `r ncol (penguins)` columns.

The mean flipper length is `r round (mean (penguins$flipper_length_mm, na.rm = TRUE), 2)`mm (NA excluded).

```{r warning=FALSE}
# load tidyverse packages
library(tidyverse)
# make a scatterplot of flipper length (y) vs bill length (x)
plot_penguins = tibble(
  x = penguins$bill_length_mm, 
  y = penguins$flipper_length_mm)
# plot flipper length (y) vs bill length (x), colored by species
ggplot(plot_penguins, aes(x = x, y = y, color = penguins$species)) + geom_point() + labs (color = "species", x = "bill length (mm)", y = "flipper length (mm)", title = "Flipper length vs Bill length" ) + theme(plot.title = element_text(hjust = 0.5))
# save the plot
ggsave("flipper_vs_bill.pdf")
```

